---
title: "Practical Machine learning"
author: "Michael"
date: "`r Sys.Date()`"
output: html_document
---

## background information


Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

## Page setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

path <- "./"
```
## import relevant R library
the following R library shall be imported for processing and analysis of data and setting working directory

```{r}
setwd("./")
path<-"./"
if(!require("tidyverse")){
  install.packages("tidyverse")
  library(tidyverse)
}
if(!require("tidyr")){
  install.packages("tidyr")
  library(tidyr)
}

if(!require("dplyr")){
  install.packages("dplyr")
  library(dplyr)
}

if(!require("ggplot2")){
  install.packages("ggplot2")
  library(ggplot2)
}
if(!require("caret")){
install.packages("caret")
 library("caret")}
if(!require("data.table")){
  install.packages("data.table")
  library("data.table")
}
# install.packages("rattle")
# install.packages("forcast")
# install.packages("pgmm")
# install.packages("rpart")
# install.packages("gbm")
# install.packages("lubridate")
# install.packages("e1071")
# install.packages("elasticnet")
```
## R Markdown
```{r}
 
if (!(file.exists("./training_data.csv"))){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = paste(path, "training_data.csv", sep = "/"),method = "curl")}
if (!(file.exists("./testing_data.csv"))){
 download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = paste(path, "test_data.csv", sep = "/"),method = "curl")}
 
 training_data <-read.csv("./training_data.csv")
 test_data <-read.csv("./test_data.csv")
```

# splitting training set
```{r}
# Now split the training into to as actual testing and validation
set.seed(1234) # Don't forget the reproducibility!
trainingDS <- createDataPartition( y = training_data$classe,
                                   p = 0.7,
                                   list = FALSE)
actual_training <- training_data[trainingDS,]
actual_validation <- training_data[-trainingDS,]

```
#Cleaning the data
we need to prepare the data for modeling. looking at the training data  there are a number of variables that have either no variance or a large fraction of missing values. we shall clean them up for a healthy modeling:
```{r}
### Now clean-up the variables w/ zero variance
### Be careful, kick out the same variables in both cases
nzv <- nearZeroVar(actual_training)
actual_training <- actual_training[,-nzv]
actual_validation <- actual_validation[,-nzv]

```
# Remove variables that are mostly NA
```{r}
mostlyNA <- sapply(actual_training,function(x) mean(is.na(x))) > 0.95
actual_training <- actual_training[,mostlyNA==FALSE]
actual_validation <- actual_validation[,mostlyNA==FALSE]

#dropping the first 5 variable as dy are not relevant predictors
actual_training <- actual_training[,-(1:5)]
actual_validation <- actual_validation[,-(1:5)]
```

## building radom forest model(rf) and a Generalized Boosted model(gbm)
```{r}
#random forest model
set.seed(111)
modelRF  <- train( classe ~.,
                   data = actual_training,
                   method = "rf",
                   trControl = trainControl(method="cv",number=3) )

set.seed(111)
modelBM <- train( classe ~.,
                  data = actual_training,
                  method = "gbm",
                  trControl = trainControl(method="repeatedcv",number = 5,repeats = 1),
                  verbose = FALSE)

# confusion matrix computation for rf model
prediction_validation_rf <- predict(modelRF,actual_validation)
confusion_matrix_rf <- confusionMatrix(prediction_validation_rf,as.factor(actual_validation$classe))
print(confusion_matrix_rf)
# confusion matrix computation for gbm model
prediction_validation_gbm <- predict(modelBM,actual_validation)
confusion_matrix_rf <- confusionMatrix(prediction_validation_gbm,as.factor(actual_validation$classe))
print(confusion_matrix_rf)

```

### predicting the test cases


You can also embed plots, for example:

```{r pressure, echo=FALSE}
# Now get the prediction in the testing portion and see what we get
prediction_testing_rf <- predict(modelRF,test_data)
print(prediction_testing_rf)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
