---
title: "Practical Machine learning"
author: "Michael"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

path <- "./"
```
## import relevant R library
the following R library shall be imported for processing and analysis of data and setting working directory

```{r}
setwd("./")
path<-"./"
if(!require("tidyverse")){
  install.packages("tidyverse")
  library(tidyverse)
}
if(!require("tidyr")){
  install.packages("tidyr")
  library(tidyr)
}

if(!require("dplyr")){
  install.packages("dplyr")
  library(dplyr)
}

if(!require("ggplot2")){
  install.packages("ggplot2")
  library(ggplot2)
}
if(!require("caret")){
install.packages("caret")
 library("caret")}
if(!require("data.table")){
  install.packages("data.table")
  library("data.table")
}
# install.packages("rattle")
# install.packages("forcast")
# install.packages("pgmm")
# install.packages("rpart")
# install.packages("gbm")
# install.packages("lubridate")
# install.packages("e1071")
# install.packages("elasticnet")
```
## R Markdown
```{r}
 
if (!(file.exists("./training.csv"))){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = paste(path, "training_data.csv", sep = "/"),method = "curl")}
if (!(file.exists("./testing.csv"))){
 download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = paste(path, "test_data.csv", sep = "/"),method = "curl")}
 
 training_data <-read.csv("./training.csv")
 test_data <-read.csv("./training.csv")
```

# splitting training set
```{r}
# Now split the training into to as actual testing and validation
set.seed(1234) # Don't forget the reproducibility!
trainingDS <- createDataPartition( y = training_data$classe,
                                   p = 0.7,
                                   list = FALSE)
actual_training <- training_data[trainingDS,]
actual_validation <- training_data[-trainingDS,]

```
#Cleaning the data
we need to prepare the data for modeling. looking at the training data  there are a number of variables that have either no variance or a large fraction of missing values. we shall clean them up for a healthy modeling:
```{r}
### Now clean-up the variables w/ zero variance
### Be careful, kick out the same variables in both cases
nzv <- nearZeroVar(actual_training)
actual_training <- actual_training[,-nzv]
actual_validation <- actual_validation[,-nzv]

```
# Remove variables that are mostly NA
```{r}
mostlyNA <- sapply(actual_training,function(x) mean(is.na(x))) > 0.95
actual_training <- actual_training[,mostlyNA==FALSE]
actual_validation <- actual_validation[,mostlyNA==FALSE]

#dropping the first 5 variable as dy are not relevant predictors
actual_training <- actual_training[,-(1:5)]
actual_validation <- actual_validation[,-(1:5)]
```

## building radom forest model(rf) and a Generalized Boosted model(gbm)
```{r}
#random forest model
set.seed(111)
modelRF  <- train( classe ~.,
                   data = actual_training,
                   method = "rf",
                   trControl = trainControl(method="cv",number=3) )

set.seed(111)
modelBM <- train( classe ~.,
                  data = actual_training,
                  method = "gbm",
                  trControl = trainControl(method="repeatedcv",number = 5,repeats = 1),
                  verbose = FALSE)

# confusion matrix computation for rf model
prediction_validation_rf <- predict(modelRF,actual_validation)
confusion_matrix_rf <- confusionMatrix(prediction_validation_rf,as.factor(actual_validation$classe))
print(confusion_matrix_rf)
# confusion matrix computation for gbm model
prediction_validation_gbm <- predict(modelBM,actual_validation)
confusion_matrix_rf <- confusionMatrix(prediction_validation_gbm,as.factor(actual_validation$classe))
print(confusion_matrix_rf)

```
You can also embed plots, for example:

```{r pressure, echo=FALSE}
# Now get the prediction in the testing portion and see what we get
prediction_testing_rf <- predict(modelRF,test_data)
print(prediction_testing_rf)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
